Extended Data

### 2021-10-05 doc/manuscript/Dow_manuscript_Nature.Rmd
big-picture rearrangement of intro

@@

This would be concerning in that the C sink of temperate deciduous forests, including those in eastern North America, is important for reducing atmospheric CO~2~ concentrations(**REFS**).

As Earth's atmospheric CO~2~ concentrations and temperatures increase, the responses and feedbacks of forests will play an important yet uncertain role in influencing atmospheric CO~2~ and climate change[@arora_carbon_2020; @friedlingstein_global_2020]. 
Understanding how trees are responding to warming temperatures and rising CO2 levels, and how growth may become limited because of interactions between the two is essential to improve our predictions. 
Under controlled conditions, spring warming increases biomass accumulation in trees, with a larger proportion of root growth compared to shoot growth[@zohner_how_2021]. 
To our knowledge, the link between stem phenology and growth has never been studied in-situ.

We know very little about how the longest-lived component of fixed carbon in trees, the woody growth, is responding to warming spring temperatures. 
In fact, we know little about the stem-growth phenology of temperate deciduous species, with multi-year records of growth phenology having been published for only a few site [@doranngeville_radial_inreview; @martin_hydrometeorology_2017]. *(We need to make sure we're not missing anything. See European study cited in A-T et al in review)* 
Tree-ring records, which can be used to examine correlations of annual growth to monthly temperatures but not to understand growth phenology, reveal that growth of temperate deciduous trees tends to be most sensitive to summer temperatures (**MORE REFS--preferably some that are already cited**), with some hints that warmer springs may have a modest positive effect on growth (e.g., Refs.[@helcoski_growing_2019] [**MORE REFS--but only those that are already cited**]).
Thus, tree-ring records do not necessarily support the suggestion that warming spring temperatures would increase annual CO~2~ sequestration. 
Characterizing phenological responses of stem growth to warming spring temperatures is critical to bridging this conceptual disconnect and understanding how forest biomass growth is likely to change as the climate warms.(Loic suggests adding numbers to illustrate the importance of stem growth as a carbon sink.)

@@

>>>>

As Earth's atmospheric CO~2~ concentrations and temperatures increase, the responses and feedbacks of forests will play an important yet uncertain role in influencing atmospheric CO~2~ and climate change[@arora_carbon_2020; @friedlingstein_global_2020]. 
Temperate deciduous forests, including those in eastern North America, are important for reducing atmospheric CO~2~ concentrations(**REFS**).
*yet models remain uncertain.*

@@

We know very little about how the longest-lived component of fixed carbon in trees, the woody growth, is responding to warming spring temperatures. 
In fact, we know little about the stem-growth phenology of temperate deciduous species, with multi-year records of growth phenology having been published for only a few site [@doranngeville_radial_inreview; @martin_hydrometeorology_2017]. *(We need to make sure we're not missing anything. See European study cited in A-T et al in review)* 
To our knowledge, the link between stem phenology and growth has never been studied *in-situ*.
Tree-ring records, which can be used to examine correlations of annual growth to monthly temperatures but not to understand growth phenology, reveal that growth of temperate deciduous trees tends to be most sensitive to summer temperatures (**MORE REFS--preferably some that are already cited**), with some hints that warmer springs may have a modest positive effect on growth (e.g., Refs.[@helcoski_growing_2019] [**MORE REFS--but only those that are already cited**]).
Thus, tree-ring records do not necessarily support the suggestion that warming spring temperatures would increase annual CO~2~ sequestration. 
Characterizing phenological responses of stem growth to warming spring temperatures is critical to bridging this conceptual disconnect and understanding how forest biomass growth is likely to change as the climate warms.(Loic suggests adding numbers to illustrate the importance of stem growth as a carbon sink.)

@@

Under controlled conditions, spring warming increases biomass accumulation in trees, with a larger proportion of root growth compared to shoot growth[@zohner_how_2021].
